\chapter{\label{chap:legacyAnalysis}From Legacy Prototypes to Unified Platform}

% ============================================================================
% PURPOSE: Tell the full transformation story in one place.
% Part I (§4.1–4.3): Analyze the legacy systems and derive requirements.
% Part II (§4.4–4.7): Describe the refactoring process and key engineering
% decisions that produced the current system.
%
% Rule: Do NOT describe the final architecture here — that is Chapter 5.
% Focus on the PROCESS and the DECISIONS, not the end result.
% TARGET: ~18-25 pages
% ============================================================================
%
% PAGE ESTIMATES: §4.1–4.3 (8–10 pages), §4.4 (3–4 pages),
%   §4.5 (2–3 pages), §4.6 (3–4 pages), §4.7 (2–3 pages)
% Total target: 18–24 pages

\section{Overview of Legacy Prototypes}
\label{sec:legacy-overview}

The Twin2MultiCloud platform builds on two independent bachelor thesis prototypes: a cost optimizer that estimated monthly hosting costs for digital twin deployments, % TODO: CITE --- Gritsch et al., Twin2Clouds (MODELS 2025)
and a cloud deployer that provisioned the required \gls{aws} infrastructure. % TODO: CITE --- Bachelor thesis (source to be added)
Neither project was designed for integration with the other. The optimizer ran entirely in the browser with no server component, while the deployer operated as a local command-line tool. This section describes each prototype's original state, focusing on the architectural decisions and technical constraints that later drove the refactoring effort documented in Sections~\ref{sec:refactoring-optimizer}--\ref{sec:refactoring-management}.

\subsection{Project A: The Cost Optimizer (``The Brain'')}
\label{sec:legacy-optimizer}

The original cost optimizer was a client-side JavaScript web application, developed as part of a conference paper submission. % TODO: CITE --- Gritsch et al., Twin2Clouds (MODELS 2025)
It consisted of 15 files totaling approximately 2,160 lines of code, with no server component. Users opened an HTML page served via Python's built-in HTTP server (\texttt{python3~-m~http.server}), filled in 12 input fields describing their \gls{iot} deployment scenario (device count, sending interval, message size, storage durations, dashboard configuration, and an optional 3D model toggle), and pressed a button to trigger the cost calculation. The result was rendered as a set of HTML cards injected directly into the page's \gls{dom}.

The calculation itself lived in a single function, \texttt{calculateCosts()}, spanning 491~lines in \texttt{cost\_calculation.js}. This function handled input parsing, validation, cost computation for all layers, cross-provider transfer cost estimation, and output rendering in one monolithic block. Listing~\ref{lst:legacy-calc-header} shows the function's opening section, where twelve form values are read from the \gls{dom} using manual \texttt{parseInt()} and \texttt{parseFloat()} calls with no type safety beyond runtime \texttt{isNaN()} checks.

\begin{lstlisting}[language=JavaScript, caption={Opening of the legacy \texttt{calculateCosts()} function: 12 DOM reads with manual type coercion.}, label={lst:legacy-calc-header}]
async function calculateCosts() {
  if (!pricing) { await loadPricingData(); }
  const numberOfDevices = parseInt(
    document.getElementById("devices").value);
  const deviceSendingIntervalInMinutes = parseFloat(
    document.getElementById("interval").value);
  const averageSizeOfMessageInKb = parseFloat(
    document.getElementById("messageSize").value);
  const hotStorageDurationInMonths = parseInt(
    document.getElementById("hotStorageDurationInMonths").value);
  // ... 8 more DOM reads, then inline validation,
  // then all layer calculations, then HTML assembly
}
\end{lstlisting}

Pricing data was stored in a static \texttt{pricing.json} file containing 115~lines of hardcoded values for \gls{aws} and Azure services. There was no mechanism to fetch live prices from cloud provider \glspl{api}; the values were accurate only at the time the file was last manually updated. Given that cloud providers adjust their pricing periodically, these figures would grow stale within weeks.

The optimizer supported only two providers: \gls{aws} and Azure. It used a graph-based approach to find the cheapest storage path: the \texttt{provider\_decision.js} module (126~lines) constructed a weighted directed graph of all possible hot$\to$cool$\to$archive storage combinations across the two providers, then ran Dijkstra's shortest-path algorithm to select the cost-minimal route. Cross-cloud data transfer was modeled through twelve explicit transfer cost functions in \texttt{data\_transfer.js} (148~lines), each computing the egress fee for a specific provider-to-provider boundary. There was no concept of a glue layer (L0) or relay functions; transfer overhead was treated as a flat monetary addition rather than a deployable infrastructure component.

Several design limitations are worth highlighting. First, the application used a reversed layer numbering compared to the current system: Layer~2 referred to storage and Layer~3 to processing, the opposite of the convention adopted in the unified platform. Second, the entire codebase was procedural JavaScript. The only class in the project was a \texttt{PriorityQueue} used by the Dijkstra implementation. There were no modules, no abstraction layers, and no separation between calculation logic and presentation. The global \texttt{pricing} variable was loaded via \texttt{fetch()} and accessed from every calculator function without encapsulation. Third, the Azure processing cost calculation contained a comment acknowledging that it reused the \gls{aws} Lambda formula verbatim: the two providers' serverless compute pricing happened to be identical at the time, so no separate implementation was written. This assumption would break if either provider changed their pricing model.

For the refactored version of the optimizer, see Chapter~\ref{chap:systemArchitecture}, Section~\ref{sec:optimizer}.

\subsection{Project B: The Cloud Deployer (``The Muscle'')}
\label{sec:legacy-deployer}

The original cloud deployer was a Python command-line application using the \texttt{boto3} \gls{aws} SDK. % TODO: CITE --- boto3 (AWS SDK for Python)
It contained 68~Python files totaling approximately 4,330~lines. The tool provisioned the entire 5-layer digital twin infrastructure on \gls{aws} through imperative SDK calls: each resource (IAM roles, Lambda functions, IoT rules, DynamoDB tables, S3 buckets, TwinMaker workspaces, Grafana instances) was created by calling the corresponding \texttt{boto3} client method directly. There was no Terraform, no CloudFormation, and no state file tracking which resources had been created. % TODO: CITE --- Terraform (HashiCorp), AWS CloudFormation

The user interacted with the deployer through a simple REPL loop. The \texttt{main.py} file (88~lines) offered five commands: \texttt{deploy}, \texttt{destroy}, \texttt{info}, \texttt{help}, and \texttt{exit}. On startup, the program initialized configuration from five JSON files and then instantiated twelve global \texttt{boto3} service clients. The \texttt{deploy} command triggered a five-phase process, invoking aggregator deployers in sequence: core infrastructure, IoT device registration, twin hierarchy setup, event action configuration, and initial value provisioning.

\begin{lstlisting}[language=Python, caption={Legacy deployer startup: 12 global client initializations and sequential deployment.}, label={lst:legacy-deployer-main}]
def main():
    globals.initialize_config()
    globals.initialize_config_iot_devices()
    globals.initialize_config_credentials()
    globals.initialize_config_events()
    globals.initialize_config_hierarchy()
    globals.initialize_aws_iam_client()
    globals.initialize_aws_lambda_client()
    # ... 10 more client initializations
    while True:
      command = input(">>> ").strip()
      if command == "deploy":
        sanity_checker.check()
        deployers.aws.core.all.AllDeployer().deploy()
        deployers.aws.iot.all.AllDeployer().deploy()
        # ... 3 more sequential phases
\end{lstlisting}

The most significant architectural issue was the \texttt{globals.py} module (257~lines), which served as a shared mutable state container for the entire application. It held twelve \texttt{boto3} client objects as module-level variables, initialized through individual \texttt{initialize\_*} functions that each used the \texttt{global} keyword. Alongside the clients, the file contained over thirty naming functions that constructed resource identifiers via string concatenation (for example, \texttt{config["digital\_twin\_name"] + "-dispatcher"}). Every other module in the codebase imported \texttt{globals} directly to access these shared clients and naming utilities. This design made it impossible to write isolated unit tests, since any test that imported a deployer module would transitively trigger \texttt{boto3} client construction, which required valid \gls{aws} credentials.

Despite this coupling, the deployer did exhibit one useful structural pattern. An abstract base class, \texttt{Deployer}, defined four methods: \texttt{deploy()}, \texttt{destroy()}, \texttt{info()}, and \texttt{log()}. Each of the 49~resource deployer files under \texttt{deployers/aws/} implemented this interface for a single cloud resource. Aggregator classes then composed these per-resource deployers into layer-level deployment units. While this composition approach provided a clear deployment ordering, it was tightly bound to \gls{aws}: the class hierarchy lived under \texttt{deployers/aws/}, and adding a second provider would have required duplicating the entire directory structure with no shared abstraction between providers.

The deployer's configuration format included a \texttt{config\_providers.json} file that mapped each architectural layer to a cloud provider:

\begin{lstlisting}[language=json, caption={Legacy provider assignment: per-layer configuration with only AWS implemented.}, label={lst:legacy-providers}]
{
  "layer_1_provider": "aws",
  "layer_2_provider": "aws",
  "layer_3_provider": "aws",
  "layer_4_provider": "aws",
  "layer_5_provider": "aws"
}
\end{lstlisting}

Although the configuration schema allowed per-layer provider selection, only \gls{aws} implementations existed. Setting any layer to \texttt{"azure"} or another provider would have been silently ignored or would have caused an import error, since no corresponding deployer classes had been written.

Additional limitations compounded the architectural constraints. Cloud credentials were stored in plaintext in \texttt{config\_credentials.json} on disk, with no encryption, no validation before deployment, and no separation of concerns between credential management and infrastructure provisioning. If a deployment failed midway, for example because an IAM role creation timed out, the deployer had no rollback mechanism: resources created before the failure remained orphaned on \gls{aws}, requiring manual cleanup through the console. There was no streaming of progress information; the only feedback during a deployment was whatever \texttt{print()} statements the individual deployer classes emitted to the terminal. The project contained no tests of any kind. Furthermore, validation was minimal: the only input check was a regular expression constraining the digital twin name to ten alphanumeric characters with hyphens and underscores.

For the refactored version of the deployer, see Chapter~\ref{chap:systemArchitecture}, Section~\ref{sec:deployer}.

\section{Gap Analysis}
\label{sec:legacy-gap-analysis}
% Systematically identify what was missing to build a production-quality
% multi-cloud Digital Twin platform.
%
% NOTE: Include a condensed side-by-side comparison table here (or at end of
% Section 4.1) showing ~15 key dimensions: language, execution model, cloud
% providers, IaC, state management, design patterns, testing, etc.
% Source data: docs/codebase_investigation_findings.md §11 (comparative analysis).
% Keep it focused on motivating the gaps — the Discussion chapter (§7.2)
% will circle back with quantitative before/after metrics.

\subsection{Lack of Integration}
% The two projects had:
% - No common data format (Optimizer output could not feed into Deployer input)
%   - Optimizer: HTML template literals injected into DOM
%   - Deployer: reads from 5 separate JSON config files
% - No shared API (Optimizer: browser-only; Deployer: CLI REPL)
% - No shared configuration model (each had its own incompatible config format)
% - No way to pass the Optimizer's "cheapest path" decision to the Deployer
%   (Dijkstra result was rendered as HTML, not structured data)
%
% DIAGRAM: Show the disconnected state — two boxes with no arrow between them.

\subsection{Absence of State Management}
% Neither project tracked state:
% - The Optimizer: no persistence of calculation results or pricing data
% - The Deployer: no record of what was deployed; "deploy and forget"
% - No way to know: "Is my twin currently deployed? On which providers?"
% - No deployment lifecycle (draft → deploying → deployed → destroying → destroyed)
%
% This made it impossible to build UX around deployment management.

\subsection{Usability Barriers}
% Both projects required deep technical knowledge to operate:
% - Manual JSON editing for all configuration
% - CLI-only interaction, no visual feedback
% - No validation: typos in config files caused cryptic errors at deploy time
% - No credential validation before attempting deployment
% - No progress feedback during (potentially 10+ minute) deployments
%
% These barriers made the prototypes unsuitable for any user beyond the authors.

\subsection{Inflexibility and Hardcoded Values}
% Hardcoded elements that prevented real-world usage:
% - Fixed AWS regions (e.g., "eu-central-1" embedded in code)
% - Fixed resource naming (collisions when deploying multiple twins)
% - Fixed pricing values (stale within weeks)
% - No support for optional architectural layers (all-or-nothing deployment)
%
% EVIDENCE: A table showing hardcoded vs. configurable elements
% in the legacy code vs. the new system.

\section{Requirements for a Unified Platform}
\label{sec:legacy-requirements}
% Derived from the gap analysis, state what the new system must provide.
% These requirements directly motivate the architecture in
% Chapter~\ref{chap:systemArchitecture}.

\subsection{Architectural Requirements}
% - AR-1: API-first design — all components must expose RESTful APIs
% - AR-2: Microservice architecture — loose coupling between Optimizer, Deployer, and Management
% - AR-3: Provider abstraction — adding a new cloud provider must not require
%   modifying existing provider implementations
% - AR-4: Infrastructure as Code — declarative, repeatable deployments
% - AR-5: Real-time feedback — deployment progress must be streamed to the user

\subsection{Functional Requirements}
% - FR-1: Persist twin configurations and deployment state in a database
% - FR-2: Fetch live pricing data from cloud provider APIs
% - FR-3: Calculate optimal cross-cloud configurations automatically
% - FR-4: Deploy and destroy infrastructure across AWS, Azure, and GCP
% - FR-5: Provide a graphical user interface for non-technical users
% - FR-6: Validate credentials, configurations, and code before deployment

\subsection{Quality Requirements}
% - QR-1: Test coverage — unit tests for all calculation and deployment logic
% - QR-2: Error handling — structured exceptions with provider/layer context
% - QR-3: Type safety — protocols, enums, and schema validation throughout
% - QR-4: Reproducibility — containerized environment with Docker Compose

% ===================== PART II: REFACTORING =====================

\section{Refactoring Methodology}
\label{sec:refactoring-methodology}
% Describe the iterative approach that was applied to both legacy systems.
%
% GUIDANCE: Explain shared patterns (Protocol, Factory, Registry) ONCE here
% with one illustrative code example. §4.5–4.7 should back-reference this
% section and focus only on what was UNIQUE to each subsystem's refactoring.
%
% Why incremental patching failed:
% - Adding Azure alongside AWS revealed hardcoded assumptions everywhere
% - No abstraction layers: provider-specific logic entangled with orchestration
% - Conditional branches for each provider would have tripled the code
%
% The iterative cycle that emerged:
% 1. Extract an interface (Python Protocol) to define a provider-agnostic contract
% 2. Refactor existing AWS code to satisfy the new Protocol
% 3. Implement the same Protocol for Azure and GCP
% 4. Wire through a factory or registry for runtime provider selection
% 5. Discover the next piece of coupled code → return to step 1
%
% This cycle repeated multiple times for both Optimizer and Deployer.

\section{Refactoring the Cost Optimizer}
\label{sec:refactoring-optimizer}
% The transformation from client-side JS app to Python FastAPI service.
% SOURCE: deprecated/optimizer/ → 2-twin2clouds/
%
% Focus on the PROCESS and KEY DECISIONS, not the final architecture:
% - Rewriting calculations from DOM manipulation to Pydantic-validated parameters
% - Replacing static pricing.json with live API fetchers (PriceFetcher Protocol)
% - Normalizing incomparable pricing dimensions across three providers
%   (per-message vs. per-hub-unit vs. per-data-volume)
% - Restructuring the monolithic calculateCosts() (491 lines) into
%   component calculators (ResourceCalculator Protocol)
% - Adding GCP as the third provider, which validated the abstraction
%
% Include 1-2 code snippets showing key Protocol definitions as evidence.
% Cross-reference Chapter~\ref{chap:systemArchitecture} for the full
% architecture description of the final system.

\section{Refactoring the Cloud Deployer}
\label{sec:refactoring-deployer}
% The transformation from imperative boto3 CLI to Terraform-based multi-provider service.
% SOURCE: deprecated/deployer/ → 3-cloud-deployer/
%
% Focus on the PROCESS and KEY DECISIONS:
% - Why imperative SDK calls could not scale to multi-cloud
%   (dependency ordering, state tracking, error recovery, no rollback)
% - The migration to Terraform: how declarative IaC solved these problems
% - Extracting CloudProvider and DeployerStrategy Protocols from the
%   existing Deployer ABC base class
% - Replacing globals.py (12 global boto3 clients) with
%   DeploymentContext (explicit dependency injection)
% - The provider registry and auto-registration pattern
% - Adding Azure and GCP providers against the established interfaces
% - Terraform file organization: one file per (provider, layer/concern)
%
% Include 1-2 code snippets. Cross-reference Chapter~\ref{chap:systemArchitecture}
% for full architecture.

\section{Building the Management API and UI}
\label{sec:refactoring-management}
% This is NEW WORK, not refactoring — there was no legacy orchestrator.
% SOURCE: twin2multicloud_backend/ + twin2multicloud_flutter/
%
% Focus on KEY ENGINEERING DECISIONS:
% - Why a separate orchestrator was needed: credential decryption, state
%   tracking, unified API for the UI. Neither Brain nor Muscle should
%   know about each other.
% - The ZIP assembly approach: converting database records into the
%   Deployer's expected file/directory format
% - SSE streaming chain: Terraform → Deployer → Management API → Flutter UI
% - Flutter architecture choices: BLoC for complex flows, Riverpod for DI
% - JWT authentication with DEBUG-mode bypass for development
%
% Keep brief — the orchestrator's architecture is fully described in
% Chapter~\ref{chap:systemArchitecture}. This section explains WHY it
% was built the way it was.
